### 2016-06-01 resubmission_materials/reviewer_response.md

Dear Dr. Markel,

Thank you for reviewing our manuscript "Ten Simple Rules for Digital Data Storage" and considering our resubmission. We appreciate the thoughtful feedback reviewers have provided and incorporated their comments into our revised manuscript. Below is a detailed accounting of our changes, with reviewer submissions highlighted in bold.

**_Reviewer #1:_** _The manuscript enumerates a set of "rules" that promote up-front thought during experimental design regarding the long term storage of data generated. Each of the ten rules has a specific focus on one aspect that researchers should consider. The authors clearly communicate the rules with relevant supporting observations.
A few minor issues:_

**Rules 4 and 5 seem like variations on the same rule -- keeping the data accessible. It would be helpful to differentiate these rules a bit more.**

**Line 21: is a good advice => is good advice**

We have fixed this typo.

Line 136: States that an example of machine readable open format that would not be easy to process include Microsoft word, but in the preceding rule (Rule 4) it states that Microsoft excel is not an open format. Perhaps just keep PDF as the example.

Lines 140-144: discussing source code doesn't really fit in this section. Perhaps move these lines to the section about metadata (Rule 7).

**Line 147: variable is a column => variable as a column**

We have fixed this typo.

**Line 147-148: unit is a table => unit as a table**

We have fixed this typo.

**Line 161: standards-compliant data, it _is_ easier to build**

We have fixed this typo.

**Line 200: meta-data in in well => meta-data in well**

We have fixed this typo

**Line 201: Extraneous close paren ) after JSON**

We have deleted this dangling parentheses.

**Line 244: Seems like an appeal to authority. Replace explicit reference to Kristen Briney**

**Line 245: Include a caution with respect to Rule 8 (ensure that the offsite "cloud" storage is as secure as the primary storage)**

**Line 311: include oxford comma "animals, fungi, and microbes"**

An oxford comma has been added.

**Line 337-338: ways to programatically query databases through**

This sentence has been restructured

**_Reviewer #1:_** _Dr. Hart and co-authors presented a comprehensive review on the essential aspects of digital data storage. While situation in reality can become very complicated, these ten simple rules can still provide good and useful guidance for researchers to develop data storage and management systems. Actually, we have pretty much followed these rules in our own practices. The manuscript was concise and well written.

Minor suggestions:_

**Page 2, Line 50: "can help" -> "can help to"**

We have have added the word to

**Page 3, Line 89: "can be difficult to impossible for others" -> "can be difficult for others"**

We have deleted the word "impossible"

**_Reviewer #3:_** This manuscript discusses ten simple rules for digital data storage. The ten rules provide simple guidelines for how researchers across a variety of disciplines can store their data for long-term access and ensure that the data are usable by others in the future. Compared with other "ten Simple Rules" papers on scientific data management and curation, this manuscript further addresses another facet of data, longer term storage best practices. The manuscript represents the collaboration among co-authors from a wide variety of academic backgrounds, who are all part of the Software Carpentry project.

**The manuscript could benefit from addressing specific needs for the storage of biological and biomedical data to be more in scope with the main themes of the journal.**

### 2016-09-01 manuscript/manuscript.tex
rebuilt manuscript, removed tex from ignore file for final submission

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

@@

@@

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

@@

\begin{flushleft}
{\Large
\textbf{Ten Simple Rules for Data Storage}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Edmund Hart $^{1 ,\ast}$,
Pauline Barmby $^{2 }$,
Jeffrey Hollister $^{3 }$,
\\
\bigskip
\bf{1} Univeristy of Vermont, Department of Biology, Burlington
\\
\bf{2} University of Western Ontario, Department of Physics and Astronomy
\\
\bf{3} US Environmental Protection Agency, Atlantic Ecology Division
\\
\bigskip
 $\ast$ E-mail: emh@emhart.info

@@

\section*{Abstract}\label{abstract}
\addcontentsline{toc}{section}{Abstract}

@@

Some example text with a citation {[}1{]}

\section*{Rule 1: Rule}\label{rule-1-rule}
\addcontentsline{toc}{section}{Rule 1: Rule}

\section*{Rule 2: Rule}\label{rule-2-rule}
\addcontentsline{toc}{section}{Rule 2: Rule}

\section*{Rule 3: Rule}\label{rule-3-rule}
\addcontentsline{toc}{section}{Rule 3: Rule}

\section*{Rule 4: Rule}\label{rule-4-rule}
\addcontentsline{toc}{section}{Rule 4: Rule}

\section*{Rule 5: Rule}\label{rule-5-rule}
\addcontentsline{toc}{section}{Rule 5: Rule}

\section*{Rule 6: Rule}\label{rule-6-rule}
\addcontentsline{toc}{section}{Rule 6: Rule}

\section*{Rule 7: Rule}\label{rule-7-rule}
\addcontentsline{toc}{section}{Rule 7: Rule}

\section*{Rule 8: Rule}\label{rule-8-rule}
\addcontentsline{toc}{section}{Rule 8: Rule}

\section*{Rule 9: Rule}\label{rule-9-rule}
\addcontentsline{toc}{section}{Rule 9: Rule}

\section*{Rule 10: Rule}\label{rule-10-rule}
\addcontentsline{toc}{section}{Rule 10: Rule}

\section*{Figure Legends}\label{figure-legends}
\addcontentsline{toc}{section}{Figure Legends}

Figures here: Will need to figure out numbering\ldots{}

\section*{Tables}\label{tables}
\addcontentsline{toc}{section}{Tables}

Tables here: Will need to figure out numbering\ldots{}

\nolinenumbers

>>>>

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref}

@@

%to handle url breaks in references
\usepackage{url}
\usepackage{hyperref}
\usepackage{breakurl}

@@

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}
% The below needed to build markdown documents containing lists with pandoc
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}
}

@@

\begin{flushleft}
{\Large
\textbf{Ten Simple Rules for Digital Data Storage}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Edmund Hart $^{1 ,\ast}$,
Pauline Barmby $^{2 }$,
David LeBauer $^{3 }$,
François Michonneau $^{4,5 }$,
Sarah Mount $^{6 }$,
Patrick Mulrooney $^{11 }$,
Timothée Poisot $^{7 }$,
Kara H. Woo $^{8 }$,
Naupaka B. Zimmerman $^{9 }$,
Jeffrey Hollister $^{10 }$\\
\bigskip
\bf{1} University of Vermont, Department of Biology, Burlington
\\
\bf{2} University of Western Ontario, Department of Physics and Astronomy
\\
\bf{3} University of Illinois at Urbana-Champaign, National Center for Supercomputing Applications and Institute for
Genomic Biology
\\
\bf{4} University of Florida, iDigBio, Florida Museum of Natural History
\\
\bf{5} University of Florida, Whitney Laboratory for Marine Bioscience
\\
\bf{6} King's College London, Department of Informatics
\\
\bf{7} Université de Montréal, Département de Sciences Biologiques
\\
\bf{8} Washington State University, Center for Environmental Research, Education, and Outreach
\\
\bf{9} University of Arizona, School of Plant Sciences
\\
\bf{10} US Environmental Protection Agency, Atlantic Ecology Division
\\
\bf{11} University of California at San Diego, San Diego Supercomputer Center
\\
\bigskip
 $\ast$ E-mail: edmund.m.hart@gmail.com

@@

@@

Data is the central currency of science, but the nature of scientific
data has changed dramatically with the rapid pace of technology. This
change has led to the development of a wide variety of data formats,
dataset sizes, data complexity, data use cases, and data sharing
practices. Improvements in high throughput DNA sequencing, sustained
institutional support for large sensor networks {[}1,2{]}, and sky
surveys with large-format digital cameras {[}3{]} have created massive
quantities of data. At the same time, the combination of increasingly
diverse research teams {[}4{]} and data aggregation in portals (e.g.~for
biodiversity data, \href{http://gbif.org/}{GBIF.org} or
\href{https://www.idigbio.org/portal/}{iDigBio}) necessitates increased
coordination among data collectors and institutions {[}5,6{]}. As a
consequence, ``data'' can now mean anything from petabytes of
information stored in professionally-maintained databases, through
spreadsheets on a single computer, to hand-written tables in lab
notebooks on shelves. All remain important, but data curation practices
must continue to keep pace with the changes brought about by new forms
of data and new data collection and storage practices.

While much has been written about both the virtues of data sharing
{[}7,8{]} and the best practices to do so {[}9,10{]}, data storage has
received comparatively less attention. Proper storage is a prerequisite
to sharing, and indeed inadequate storage contributes to the phenomenon
of data decay or to ``data entropy'', where data, whether publicly
shared or not, becomes less accessible through time {[}11--14{]}. Best
practices for data storage often begin and end with, ``deposit your data
in a community standard repository.'' This is good advice, especially
considering your data is most likely to be reused if it is available on
a community site. Community repositories can also provide guidance for
best practices. As an example, if you are archiving sequencing data, a
repository such as those run by the National Center for Biotechnology
Information (NCBI) (e.g.~GenBank) provides not only a location for data
archival, but also encourages a set of practices related to consistent
data formatting and the inclusion of appropriate metadata. However, data
storage policies are highly variable between repositories {[}15{]}. A
data management plan utilizing best practices across all stages of the
data life cycle will facilitate transition from local storage to
repository {[}16{]}. Similarly, having such a plan can facilitate
transition from repository to repository if funding runs out or
requirements change. Good storage practices are important even (or
especially) in cases where data may not fit with an existing repository,
where only derived data products (versus raw data) are suitable for
archiving, or in the case where an existing repository may have lax
standards.

This manuscript describes 10 simple rules for digital data storage that
grew out of a long discussion among instructors for the Software and
Data Carpentry initiatives {[}17,18{]}. Software and Data Carpentry
instructors are scientists from diverse backgrounds who have encountered
a variety of data storage challenges and are active in teaching other
scientists best practices for scientific computing and data management.
Thus, this paper represents a distillation of collective experience, and
hopefully will be useful to scientists facing a variety of data storage
challenges. We additionally provide a glossary of common vocabulary for
readers who may not be familiar with particular terms.

\section*{Rule 1: Anticipate how your data will be
used}\label{rule-1-anticipate-how-your-data-will-be-used}
\addcontentsline{toc}{section}{Rule 1: Anticipate how your data will be
used}

One can avoid most of the troubles encountered during the analysis,
management, and release of data by having a clear roadmap of what to
expect \emph{before} data acquisition starts. For instance:

\begin{itemize}
\tightlist
\item
  How will the raw data be received? Are they delivered by a machine or
  software, or typed-in?
\item
  What is the format expected by the software used for analysis?
\item
  Is there a community standard format for this type of data?
\item
  How much data will be collected, and over what period of time?
\end{itemize}

The answers to these questions can range from simple cases (e.g.,
sequencing data stored in the FASTA format, which can be used ``as is''
throughout the analysis), to experimental designs involving multiple
instruments, each with its own output format and processing conventions.
Knowing the state in which the data needs to be at each step of the
analysis can help to (i) identify software tools to use in converting
between data formats, (ii) orient technological choices about how and
where the data should be stored, and (iii) rationalize the analysis
pipeline, making it more amenable to re-use {[}19{]}.

Also key is the ability to estimate the storage volume needed to store
the data, both during and after the analysis. The required strategy will
differ for datasets of varying size. Smaller datasets (e.g.~a few
megabytes in size) can be managed locally with a simple data management
plan, whereas larger datasets (e.g.~gigabytes to petabytes) will in
almost all cases require careful planning and preparation (Rule 10).

Lastly, early consideration and planning should be given to the metadata
of the project. A plan should be developed early as to what metadata
will be collected, and how it will be maintained and stored (Rule 7).
Also be sure to consider community software tools that can facilitate
metadata curation and repository submission. Examples in the biological
sciences include \emph{Morpho} for ecological metadata {[}20{]} and
\emph{mothur} {[}21{]} for submitting to NCBI's Sequence Read Archive.

\section*{Rule 2: Know your use case}\label{rule-2-know-your-use-case}
\addcontentsline{toc}{section}{Rule 2: Know your use case}

Well-identified use cases make data storage easier. Ideally, prior to
beginning data collection, researchers should be able to answer the
following questions:

\begin{itemize}
\tightlist
\item
  Should the raw data be archived (Rule 3)?
\item
  Should the data used for analysis be prepared once, or re-generated
  from the raw data each time (and what difference would this choice
  make for storage, computing requirements, and reproducibility)?
\item
  Can manual corrections be avoided in favor of programmatic or
  self-documenting approaches (e.g., Jupyter notebook or R markdown)?
\item
  How will changes to the data be tracked, and where will these tracked
  changes be logged?
\item
  Will the final data be released, and if so, in what format?
\item
  Are there restrictions or privacy concerns associated with the data
  (e.g.~for survey results with personally identifiable information
  (PII), threatened species, or confidential business information)?
\item
  Will institutional validation be required prior to releasing the data?
\item
  Does the funding agency mandate data deposition in a publicly
  available archive, and if so, when, where, and under what license?
\item
  Does the target journal mandate data deposition?
\end{itemize}

None of these questions have universal answers, nor are they the only
questions to ask before starting data acquisition. But knowing the what,
when, and how of \emph{your} use of the data will bring you close to a
reliable roadmap on how to handle data from acquisition through
publication and archival.

\section*{Rule 3: Keep raw data raw}\label{rule-3-keep-raw-data-raw}
\addcontentsline{toc}{section}{Rule 3: Keep raw data raw}

Since analytical and data processing procedures improve or otherwise
change over time, having access to the `raw' (unprocessed) data can
facilitate future re-analysis and analytical reproducibility. As
processing algorithms improve and computational power increases, new
analyses will be enabled that were not possible at the time of the
original work. If only derived data are stored, it can be difficult for
other researchers to confirm analytical results, to assess the validity
of statistical models, or to directly compare findings across studies.

Therefore, data should always be kept in raw format whenever possible
(within the constraints of technical limitations). In addition to being
the most appropriate way to ensure transparency in analysis, having the
data stored and archived in their original state gives a common point of
reference for derivative analyses. What constitutes sufficiently ``raw''
data is not always clear (e.g., ohms from a temperature sensor or images
of an Illumina sequencing flowcell are generally not archived after the
initial processing). Yet the spirit of this rule is that data should be
as ``pure'' as possible when they are stored. If derivations occur, they
should be documented by also archiving relevant code and intermediate
datasets.

A cryptographic hash (e.g., SHA or MD5) of the raw data should be
generated and distributed with the data. These hashes ensure that the
dataset has not suffered any silent corruption/manipulation while being
stored or transferred (see
\href{https://www.xsede.org/news/-/news/item/6390}{Internet2 Silent Data
Corruption}). For large enough datasets the likelihood of silent data
corruption is high. This technique has been widely used by many Linux
distributions to distribute images and has been very effective with
minimal effort.

\section*{Rule 4: Store data in open
formats}\label{rule-4-store-data-in-open-formats}
\addcontentsline{toc}{section}{Rule 4: Store data in open formats}

To maximize accessibility and long-term value, it is preferable to store
data in formats whose specifications are freely available. The
appropriate file type will depend on the data being stored (e.g.~numeric
measurements, text, images, video), but the key idea is that accessing
data should not require proprietary software, hardware, or purchase of a
commercial license. Proprietary formats change, maintaining
organizations go out of business, and changes in license fees make
access to data in proprietary formats unaffordable and risky for
end-users. Examples of open data formats include comma-separated values
(CSV) for tabular data, hierarchical data format (HDF) {[}22{]} and
NetCDF {[}23{]} for hierarchically structured scientific data, portable
network graphics (PNG) for images, KML (or other Open Geospatial
Consortium (OGC) format) for spatial data, and extensible markup
language (XML) for documents. Examples of closed formats include DWG for
AutoCAD drawings, Photoshop document (PSD) for bitmap images, Windows
Media Audio (WMA) for audio recording files, and Microsoft Excel (XLS)
for tabular data. Even if day-to-day processing uses closed formats
(e.g., due to software requirements), data being stored for archival
purposes should be stored in open formats. This is generally not
prohibitive; most closed-source software products enable users to export
data to an open format.

Not only should data be stored in an open format, but it should also be
stored in a format that computers can easily use for processing. This is
especially crucial as datasets become larger. Making data easily-usable
is best achieved by using standard data formats that have open
specifications (e.g., CSV, XML, JSON, HDF5), or by using databases. Such
data formats can be handled by a variety of programming languages, as
efficient and well-tested libraries for parsing them are typically
available. These standard data formats also ensure interoperability,
facilitate re-use, and reduce the chances of data loss or mistakes being
introduced during conversion between formats. Examples of
machine-readable open formats that would \emph{not} be easy to process
include data included in the text of a PDF file, or scanned images of
tabular data from a paper source.

\section*{Rule 5: Data should be structured for
analysis}\label{rule-5-data-should-be-structured-for-analysis}
\addcontentsline{toc}{section}{Rule 5: Data should be structured for
analysis}

To take full advantage of data, it can be useful for it to be structured
in a way that makes use, interpretation, and analysis easy. One such
structure for data stores each variable as a column, each observation as
a row, and each type of observational unit as a table (Fig. 1). The
technical term for this structure is `Codd's 3rd normal form', but it
has been made more accessible as theconcept of \emph{tidy} data
{[}24{]}. When data is organized in this way, the duplication of
information is reduced and it is easier to subset or summarize the
dataset to include the variables or observations of interest.

\textbf{Figure 1}: Example of an untidy dataset (A) and its tidy
equivalent (B). Dataset A is untidy because it mixes observational units
(species, location of observations, measurements about individuals), the
units are mixed and listed with the observations, more than one variable
is listed (both latitude and longitude for the coordinates, and genus
and species for the species names), and several formats are used in the
same column for dates and geographic coordinates. Dataset B is an
example of a tidy version of dataset A that reduces the amount of
information that is duplicated in each row, limiting chances of
introducing mistakes in the data. By having species in a separate table,
they can be identified uniquely using the Taxonomic Serial Number (TSN)
from the Integrated Taxonomic Information System (ITIS), and it makes it
easy to add information about the classification of these species. It
also allows researchers to edit the taxonomic information independently
from the table that holds the measurements about the individuals. Unique
values for each observational unit facilitate the programmatic
combination of information using ``join'' operations. With this example,
if the focus of the study for which these data were collected is based
upon the size measurements of the individuals (weight and length),
information about ``where'', ``when'', and ``what'' animals were
measured can be considered meta-data. Using the tidy format makes this
distinction clearer.

One axiom about the structure of data and code holds that one should
``write code for humans, write data for computers'' {[}25{]}. When data
can be easily imported and manipulated using familiar software, whether
via a scripting language, a spreadsheet, or any other computer program
that can import these common files, data becomes easier to re-use.
Furthermore, having the source code for the software doing the analysis
available provides provenance for how the data is processed and
analyzed. This makes analysis more transparent, since all assumptions
about the structure of the data are implicitly stated in the source
code. This also enables extraction of the analyses performed, their
reproduction, and their modification.

Interoperability is facilitated when variable names are mapped to
existing data standards. For instance, for biodiversity data, the
\href{http://www.tdwg.org/standards/450/}{Darwin Core Standard} provides
a set of terms that describe observations, specimens, samples, and
related information for a taxa. For earth science and ecosystem models
and data, the \href{https://cfconventions.org}{Climate Forecasting
Conventions} are widely adopted, such that a large ecosystem of software
and data products exist to reduce the technical burden of reformatting
and reusing large and complex data. Because each term in such standards
is clearly defined and documented, each dataset can use the terms
consistently; this facilitates data sharing across institutions,
applications, and disciplines. With machine-readable,
standards-compliant data, it becomes easier to build an Application
Programming Interface (API) to query the dataset and retrieve a subset
of interest, as outlined in Rule 10.

\section*{Rule 6: Data should be uniquely
identifiable}\label{rule-6-data-should-be-uniquely-identifiable}
\addcontentsline{toc}{section}{Rule 6: Data should be uniquely
identifiable}

To aid reproducibility, the data used in a scientific publication should
be uniquely identifiable. Ideally, datasets should have a unique
identifier such as a Digital Object Identifier (DOI), Archival Resource
Key (ARK), or a persistent URL (PURL). An increasing number of online
services, such as \href{http://figshare.com/}{Figshare},
\href{http://zenodo.org}{Zenodo}, or
\href{http://www.dataone.org}{DataOne}, are able to provide these.
Institutional initiatives also exist, and are known to your local
librarians. Some repositories may require specific identifiers and these
could change with time. For instance, NCBI sequence data will in the
future only be identified by ``accession.version'' IDs. The ``GI''
identifiers (in use since 1994) will be retired in late 2016 {[}26{]}.

Even as identifier standards may change over time, datasets can evolve
over time as well. In order to distinguish between different versions of
the same data, each dataset should have a distinct name, which includes
a version identifier. A simple way to do this is to use date stamps as
part of the dataset name. Using the ISO 8601 standard avoids regional
ambiguities: it mandates the date format \texttt{YYYY-MM-DD} (i.e.~from
largest time unit to smallest). For example, the date ``February 1st,
2015'', while written as \texttt{01-02-2015} in the UK and
\texttt{02-01-2015} in the US, is unambiguous (\texttt{2015-02-01})
under this standard.

Semantic versioning is a richer approach to solving the same problem
{[}27{]}. The CellPack datasets are an example of this {[}28{]}. A
semantic version number takes the form: \texttt{Major.Minor.Patch}, e.g.
\texttt{0.2.7}. The \emph{major version} numbers should be incremented
(or bumped) when a dataset scheme has been updated, or some other change
is made that is not compatible with previous versions of the data with
the same major version number. This means that an experiment using
version \texttt{1.0.0} of the dataset may not run on version
\texttt{2.0.0} without changes to the data analysis. The \emph{minor
version} should be bumped when a change has been made which is
compatible with older versions of the data with the same major version.
This means that any analysis that can be performed on version
\texttt{1.0.0} of the data is repeatable with version \texttt{1.1.0} of
the data. For example, adding a new year in a temporal survey will
result in a bump in the minor version. The \emph{patch version} number
is bumped when typos or bugs have been fixed. For example version
\texttt{1.0.1} of a dataset may fix a typo in version \texttt{1.0.0}.

\section*{Rule 7: Link relevant
metadata}\label{rule-7-link-relevant-metadata}
\addcontentsline{toc}{section}{Rule 7: Link relevant metadata}

Metadata is the contextual information required to interpret data (Fig.
1) and should be clearly defined and tightly integrated with data. The
importance of metadata for context, reusability, and discovery has been
written about at length in guides for data management best practices
{[}9,13,29{]}.

Metadata should be as comprehensive as possible, using standards and
conventions of a discipline, and should be machine-readable. Metadata
should always accompany a dataset, wherever it is stored, but the best
way to do this depends on the format of the data. Text files can contain
meta-data in well-defined text files such as XML or JSON. Some file
formats are self-documenting; for example NetCDF, HDF5, and many image
file formats allow for embedded metadata {[}22,23{]}. In a relational
database, metadata tables can be clearly labeled and linked to the data.
Ideally, a schema will be provided that also shows the linkages between
data tables and metadata tables. Another--simpler--scenario is a set of
flat (non-hierarchical) text files--in this case a semantically
versioned, compressed archive should be created that includes metadata.

Whatever format is used for archiving, the goal should be to make the
link between metadata and data as clear as possible. The best approach
is dependent on the archiving plan used, but even if the dataset is
archived solely for personal use, metadata will provide crucial context
for future reuse.

\section*{Rule 8: Adopt the proper privacy
protocols}\label{rule-8-adopt-the-proper-privacy-protocols}
\addcontentsline{toc}{section}{Rule 8: Adopt the proper privacy
protocols}

In datasets where privacy is important, be sure to have a plan in place
to protect data confidentiality. You should consider the different data
stakeholders when developing privacy protocols for your data storage.
These stakeholders include funding agencies, human subjects or entities,
collaborators, and yourself. Both the NSF and NIH have data sharing
policies in their grant guidelines to prevent sharing personally
identifiable information, and to anonymize data on human subjects.

In small datasets, a lookup table (protecting PII by removing it and
replacing it with a unique id that maps to the sensitive data in an
external dataset) is enough to anonymize a minimal amount of personal
information. Hashing techniques are susceptible to a number of attacks,
and all hashed data can eventually be determined. Famously, New York
City officials shared what they thought was anonymized data on cab
drivers and over 173 million cab rides. However, it was quickly
recognized that the city anonymized the data with a simple MD5 hashing
scheme and all 20 GB of data were de-anonymized in a matter of hours
{[}30{]}. This type of error can be prevented by asking a trusted
colleague or security personnel to try to ``crack'' anonymised data
before releasing it publicly. Often the person who has produced the data
is not well placed to check the fine details of their own security
procedures. If possible, the best solution is to remove any sensitive
data that is not required from the dataset prior to distribution.

In more problematic cases, the data itself allows identifiability: this
is the case with human genomic data that map directly onto a subject's
identity {[}31{]}. Methods for dealing with these complex issues at the
intersection of data storage and privacy are rapidly evolving, and
include storing changes against a reference genome to help with privacy
and reduce overall data volumes {[}32,33{]}, and/or bringing computation
to data storage facilities instead of vice versa {[}34{]}. Having a plan
for privacy before data is acquired is important, because it can
determine or limit how data will be stored.

\section*{Rule 9: Have a systematic backup
scheme}\label{rule-9-have-a-systematic-backup-scheme}
\addcontentsline{toc}{section}{Rule 9: Have a systematic backup scheme}

Every storage medium can fail, and every failure can result in loss of
data. Researchers should therefore back data up at all stages of the
research process. Data stored on local computers or institutional
servers during the collection and analysis phases of a project should be
backed up to other locations to protect against data loss. No backup
system is failsafe (see the stories of the
\href{https://www.insidehighered.com/news/2014/05/16/dedoose-crash-shows-dangers-handing-data-cloud-services}{Dedoose
crash} and the
\href{http://thenextweb.com/media/2012/05/21/how-pixars-toy-story-2-was-deleted-twice-once-by-technology-and-again-for-its-own-good/}{near
deletion of Toy Story 2}), so more than one backup system should be
used. Ideally you should have two on-site copies (such as on a computer,
an external hard drive, or a tape) and one off-site copy (e.g.~cloud
storage){[}35{]}, with care taken to ensure that the off-site copy is as
secure as the on-site copies. Keeping backups in multiple locations
additionally protects against data loss due to theft or natural
disasters.

Researchers should test their backups regularly to ensure that they are
functioning properly. Common reasons for backup failure include:

\begin{itemize}
\tightlist
\item
  faulty backup software
\item
  incorrect configuration (e.g., not backing up sub-directories)
\item
  encryption (e.g., someone encrypted the backups but later lost the
  password to decrypt them)
\item
  media errors
\end{itemize}

Consider the backup plan of your selected data repository before
publishing your data and if possible, find out about the long-term
storage plans of the repository. Many repositories mirror the data they
host on multiple machines. Are there plans in place to keep data
available if the organization that manages the repository dissolves?

\section*{Rule 10: The location and method of data storage depend on how
much data you
have}\label{rule-10-the-location-and-method-of-data-storage-depend-on-how-much-data-you-have}
\addcontentsline{toc}{section}{Rule 10: The location and method of data
storage depend on how much data you have}

The storage method you should choose depends on the size and nature of
your data, the cost of storage and access over time, the time it takes
to transfer the data, how the data will be used, and any privacy
concerns. Data is increasingly generated in the range of many terabytes
by environmental sensors, satellites, automated analytical tools,
simulation models, and nucleic acid sequencers. Even larger data
generating machines, like the Large Hadron Collider (LHC) and the Large
Scale Synoptic Survey Telescope (LSST), generate many terabytes (TB) per
day, rapidly accumulating to petabyte (PB) scale over the course of any
particular study. While the cost of storage continues to decrease, the
volume of data to be stored impacts the choice of storage methods and
locations: for large datasets it is necessary to balance the cost of
storage with the time of access and costs of re-generating the data.
With new commercial cloud offerings (e.g., Amazon S3) the cost of
retrieving the data might exceed the cost of analysis or of
re-generating the data from scratch.

When data takes too long to transfer or is costly to store, it can
become more efficient to use a computer system for analysis that can
directly access and use the data in place instead of first transferring
it to a local machine. Inactive data can be put in longer-term storage;
this is less expensive, but can take longer to retrieve. Some storage
systems automatically migrate `stale' files to longer term storage.
Alternatively, some computing can be done `in the database' or `on disk'
via database query languages (e.g.~SQL, MapReduce) that perform basic
arithmetic, or via the use of procedural languages (e.g.~R, Python, C)
embedded in the database server. Modern database technologies such as
HDFS and Spark allow these computations to be done on data of almost any
size. When data is larger than locally available RAM, it can be handled
by conducting analyses on a `big memory' node, which most
high-performance computing centers have deployed. Relying on tight
software/hardware integration, these can allow for the analysis of
datasets around 1-4 TB in size. This allows the user to read in and use
a large dataset without special tools.

If you regularly only need access to a small subset of your data or need
to share it with many collaborators, a web-based API (Application
Programming Interface) might be a good solution. Using this method, many
users can send requests to a web service which can subset the data,
perform in-database computation, and return smaller volumes of data as
specific slices. Tools based on web services make it easier to find and
download data, and facilitate analysis via reproducible scripts.
However, they can also lead to excessive and careless abuse of resources
without proper safeguards in place. The time required to re-download and
recompute results can be reduced by `caching'. Caching stores copies of
downloads and generated files that are recognized when the same script
is run multiple times.

\section*{Further Reading and
Resources}\label{further-reading-and-resources}
\addcontentsline{toc}{section}{Further Reading and Resources}

Digital data storage is a vast topic; the references given here and
elsewhere in this paper provide some starting points for interested
readers. For beginning users of scientific data,
\href{http://datacarpentry.org}{Data Carpentry} offers workshops and
resources on data management and analysis, as do the DataONE education
modules {[}36{]}. For librarians and others who are responsible for data
archiving, Data Curation Profiles {[}37{]} may be of interest.

\section*{Glossary used in the
manuscript}\label{glossary-used-in-the-manuscript}
\addcontentsline{toc}{section}{Glossary used in the manuscript}

\subsection*{Projects and Initiatives}\label{projects-and-initiatives}
\addcontentsline{toc}{subsection}{Projects and Initiatives}

\begin{itemize}
\item
  \textbf{Global Biodiversity Information Facility} (GBIF,
  \url{http://www.gbif.org}) provides an international open data
  infrastructure to publish and disseminate biodiversity information.
\item
  \textbf{Integrated Digitized Biocollections} (iDigBio,
  \url{https://www.idigbio.org}) is a project funded by the National
  Science Foundation that facilitates the digitization of natural
  history collections, and provides data and images for biological
  specimens.
\item
  \textbf{Integrated Taxonomic Information System} (ITIS,
  \url{http://www.itis.gov}) is an international partnership of
  governmental organizations that aims at providing authoritative
  taxonomic information for plants, animals, fungi, and microbes.
\end{itemize}

\subsection*{File formats}\label{file-formats}
\addcontentsline{toc}{subsection}{File formats}

\begin{itemize}
\item
  \textbf{Comma-Separated Values} (CSV) and \textbf{Tab-Separated
  Values} (TSV) are plain text file formats used to store tabular data
  where each row is represented by a line in the file, and each field
  (column) is separated by a comma (for CSV) or by the Tab character
  (for TSV).
\item
  \textbf{FASTA} is a simple and widely used file format used to
  represent sequences of nucleotides or amino-acids in plain text making
  it easy to manipulate programmatically.
\item
  \textbf{Hierarchical Data Format} (HDF5) is an open-source binary file
  format designed to store large amounts of data (and their associated
  metadata) by providing a hierarchical structure that could be compared
  to how a hard drive is organized with directories and files. It is
  maintained by the non-profit HDF Group, a spin off of the National
  Center for Supercomputing Applications (NCSA).
\item
  \textbf{JavaScript Object Notation} (JSON) is a plain text file format
  typically used to store arbitrarily structured data in the form of
  keys and values. It can be used to store non-relational databases as
  it does not rely on a tabular data format. In many respects, it has
  been replacing XML.
\item
  \textbf{Network Common Data Form} (NetCDF) is an open-source binary
  file format designed to store large datasets in array-oriented
  scientific data as typically used in the geosciences. It is maintained
  by Unidata, a non-profit member of the University Corporation for
  Atmospheric Research (UCAR) which is funded by the National Science
  Foundation.
\item
  \textbf{Extensible Markup Language} (XML) is a markup language and the
  file format used to store documents written with it. It is used to
  represent arbitrary data structures and is both human and
  machine-readable.
\end{itemize}

\subsection*{Programming and
algorithms}\label{programming-and-algorithms}
\addcontentsline{toc}{subsection}{Programming and algorithms}

\begin{itemize}
\item
  \textbf{Web Application Programming Interface} (API) provide ways to
  programmatically query databases through the internet. They notably
  allow users to retrieve and work with a small slice of a large
  dataset.
\item
  \textbf{Hadoop Distributed File System} (HDFS) is a Java-based file
  system where data is stored in small chunks across multiple redundant
  nodes.
\item
  \textbf{MapReduce} is a style of programming designed to work with
  large datasets in parallel computing environments. Such programs are
  composed of a \textbf{map} procedure where the dataset is sliced into
  several pieces, and a \textbf{reduce} procedure where summary
  operations are then applied to each of the slices.
\item
  \textbf{Secure Hash Algorithm 2} (SHA-2) is a family of Secure Hashing
  Algorithms used in cryptographic analysis, often to verify the
  integrity of a file. A cryptographic hash function converts a
  ``message'' (e.g., passwords, file content) into an encrypted value.
  Cryptographic hash functions are easy to compute from the message, but
  it should be impossible to recover the message from the output, and
  any modifications to the message should also modify the output. The
  SHA algorithms are often used in preference to similar tools such as
  MD5 (mentioned in Rule 8), which are no longer secure. All hashing
  algorithms are vulnerable to brute force attacks. Key Derivation
  Function (KDF) implementations like BCrypt and PBKDF2 are considered
  significantly more secure, but by design more costly to compute.
\item
  Apache \textbf{Spark} is an open source computing platform for
  querying large data sets in memory, in contrast to on-disk based
  methods like MapReduce.
\item
  \textbf{Structured Query Language} (SQL) is a programming language
  used to interact with relational database management systems.
\end{itemize}

\subsection*{Hardware}\label{hardware}
\addcontentsline{toc}{subsection}{Hardware}

\begin{itemize}
\tightlist
\item
  \textbf{mega-, giga-, tera-, peta-bytes} are units of digital
  information and are used to measure the size of datasets or the
  storage media. Originally a byte was the minimum amount of memory
  needed to store a single character of text in a computer. The prefixes
  mega-, giga-, tera- and peta- refer to the international system of
  units for the multiple of the unit, and correspond to \(10^{6}\),
  \(10^{9}\), \(10^{12}\), and \(10^{15}\) abbreviated M, G, T and P
  respectively.
\end{itemize}

\subsection*{Persistent identifiers}\label{persistent-identifiers}
\addcontentsline{toc}{subsection}{Persistent identifiers}

\begin{itemize}
\item
  \textbf{Archival Resource Key} (ARK) identifiers are URLs designed to
  support long-term access to information online.
\item
  \textbf{Digital Object Identifier} (DOI) provides unique and
  persistent identifiers for electronic documents (in particular journal
  articles and datasets) on the internet. The uniqueness of the
  identifiers is guaranteed by a central registry. By dissociating the
  identifier and the location of the document (i.e., the URL), the DOI
  can remain fixed even if the location of the digital object it is
  pointing to changes.
\item
  \textbf{Persistent Uniform Resource Locator} (PURL) is an URL used to
  redirect to the location of an electronic object on the internet. DOI
  and ARK are examples of implementations of PURL.
\item
  \textbf{Uniform Resource Locator} (URL) gives the location of an
  object on the World Wide Web; the most familiar type of URL is a
  website address.
\end{itemize}
